|

### 2021-10-01 appendix/appendix.Rmd
proofreading the appendix

@@

[^danilo]: Senior Lecturer, School of Social and Political Sciences, University of Lincoln, <danilofreire@gmail.com>, <https://danilofreire.github.io>.

\appendix
\addtocounter{section}{2}

@@

### Primitives of the Model

Consider an strategic interaction between a mayor ($M$) and $N \geq 3$ city councilors, indexed in $i \in \{1, 2, \cdots, N \}$. The mayor is the head of the local executive branch and has the prerogative of proposing a vector of policies voted by the city councilors. Policy proposals are a combination of public goods provision $g$, rents for the mayor $r$, and a vector of transfer for city councilors \textbf{x}.

The city council votes the mayor's proposal, and if they accept it, it is implemented. Otherwise, a reversal policy takes place. Reversal policy in this context means that the mayor's proposal failed to secure the majority of council votes. The decision process leaves the mayor's office and either immediately stops or is transferred to the council. The types of reversal policies are crucial for my argument, as it changes the relative strength of the council in their bargaining with the mayor. I investigate three types of reversal policies: a \textit{non-partisan} reversal mechanism, where city councilors have no partisan affiliation; a \textit{partisan} reversal mechanism, where the council is divided into supporters and non-supporters of the mayor; and finally, a \textit{hybrid} reversal mechanism, that combines partisan with non-partisan features. The main paper focuses on the hybrid model, as it is the more realistic scenario. Despite that, comparing the partisan and non-partisan reversal mechanisms clearly illustrates my argument that the partisan mechanism drives my results.

The mayor has to convince at least half of the city councilors to support her policy proposals. Similar to \citet{buenodemesquita2005logic}, I assume that the selectorate, i.e., the individuals that have a say in the mayor's policy proposals, are the $N$ city councilors. The winning coalition, the minimum number of members who have to be convinced to implement the policy, equals half of the selectorate $N/2$. City councilors are moved by policy concerns $p$ and transfers received from the mayor or the councilor recognized as the proposer \textbf{x}. I leave the definition of these transfers open, as this facilitates the relationship with several instruments that mayors can use to pressure city councilors. To illustrate, examples of transfers are granting electoral favors, spending in areas that the councilor has political or personal interests, giving portfolio and public jobs within the municipal bureaucracy, or even campaign contributions and bribes. The use of transfers affects the councilors' choices and generates a governing cost of $C_G$. The governing costs vary with the reversal mechanism and the legislature size.

Finally, after paying the govern costs, the mayor proposes a level of public goods provision ($g$) and rents ($r$). Public goods provision helps the mayor to get reelected. Rents are for direct mayor's consumption and do not contribute to the mayor's electoral success. This makes the mayor's expected utility a sum of the gains from rents and the benefits from reelection. Both the utilities from rents $u(.)$ and the probability of reelection $\pi(.)$ are concave functions, meaning that more rents or public goods increase the utility at a decreasing rate.\footnote{In technical terms, $u'>0$ and $\pi'>0$; and $u''<0$ and $\pi''<0$.} The probability of reelection is multiplied by the benefit from holding office $B_M > 0$. This benefit captures the tangible and intangible gains that the mayor perceives from holding the public office. I assume that the benefits are high enough to rule out an equilibrium with only rents and transfers, which is equivalent to saying that mayors strongly prefer to be reelected. If the mayor's policy proposal is approved, then the expected utility for the mayor is:

@@

The policy choices of the mayor are subject to the municipal budget constraint. The municipality has $R > 0$ resources, and cannot run debts, meaning that the budget must be balanced. I also assume that there are enough resources for the mayor to govern.

Let $C_G$ be the costs of governing. The budget balance constraint requires the offers that the mayor makes to the city councilors to satisfy the following inequality:

@@

The expected utility for the city councilors depends on the reversal policy. As discussed, I explore three different mechanisms. First, in \textit{non-partisan} reversal mechanism, when the council rejects the mayor's proposal, a reversal stage starts with resources diminish by a factor of $\delta \in (0,1)$ and the selection of one councilor randomly. The randomly selected councilor becomes the new proposer. If her proposal is accepted, it is implemented. If rejected, the budget shrinks again by a factor $\delta$, and another city councilor is recognized to propose. The process repeats until one proposal is finally accepted.

In the second reversal mode, the \textit{partisan} reversal mechanism, I assume that each councilor has a party affiliation. Party affiliations are mutually exclusive (a councilor cannot belong to two parties at once). If a city councilor is affiliated to a party aligned with the mayor, I say that she belongs to the government coalition $G \subset \{1, 2, \cdots, N \}$. Otherwise, the city councilor belongs to the opposition $O = \{1, 2, \cdots, N \} \setminus G$. The mayor's policy generates a political value of $p > 0$ for the councilors, and whether it adds or subtracts from the councilor's utility depends on the party affiliation of the councilor. If the mayor's proposal is rejected, a city councilor aligned with the mayor that should have received $p$ from the policy now receives zero. On the other hand, a city councilor that belongs to the opposition would receive $-p$ from the mayor's policy and now receive zero. Although these assumptions are strong, they capture the underlying political dynamics of partisanship: the political preferences in public goods provision are shared by councilors aligned with the mayor. Voting against the mayor's proposals would represent defying the partisan label.

Lastly, in the \textit{hybrid} reversal mechanism, I combine partisanship with legislature bargaining costs. Councilors in this mechanism receive a partisan utility from the mayor's policy, but if they reject the policy, a reversal policy similar to the non-partisan reversal mechanism takes place. In the mayor's proposal stage, city councilors that belong to the mayor's coalition receive a gain of $p > 0$ when the mayor's policy is implemented. On the other hand, opposition members receive $-p$ when the mayor's proposal is implemented. If the mayor's proposal is rejected, as the city council does not control the Executive when the two branches are separated, no public goods are provided ($p=0$). Moreover, in equilibrium, I assume that city councilors would prefer to receive transfers instead of providing public goods. This aligns with the empirical evidence in most developing economies.

@@

\begin{enumerate}
  \item The mayor learns how many government $|G|$ and opposition $|O|$ legislators were elected.
  \item The mayor proposes a policy vector $(r, g, x)_M$.
  \item The city council votes the proposal.
  \begin{itemize}
    \item If the council accepts, the policy is implemented, and the game ends. 
    \item Otherwise, the reversal policy is implemented.
  \end{itemize}
  \item (Reversal Policy) Depends on which mechanism, I propose three different reversal policies. The details are in the following sections.
\end{enumerate}

The solution concept I use in this model is the Sub-game Perfect Nash Equilibrium. A Sub-game Perfect Nash Equilibrium requires that the strategies played are a Nash Equilibrium in each of the subgames of the game. The solution is found using backward induction. In the model with infinitely repeated proposals, I extend the equilibrium concept to require that the equilibrium is Stationary. A Stationary Sub-game Perfect Equilibrium requires that, at each given point of the game, if a politician accepts an offer at time $k+1$, she should take the same offer at time $k$. Stationarity provides the intending to find a proposal that would be accepted at any stage in the game. Applying this reasoning, I can find a sequence of offers for each point $k$ in time. For the mayor, the optimal offer is at $k=0$, representing a no delay in policy implementation.

\begin{definition}
The equilibrium for the game consists of a sequence of strategic policy vectors $(g^*, r^*, x^*)_k$, indexed in $k \in \{0, 1, \cdots, \infty\}$, such that, at every stage of the game, the mayor and the councilors have no profitable deviations from the equilibrium.
\end{definition}

Moreover, I assume that the players play no dominated strategies, even when there is less than sufficient support for implementing the policy.

@@

Solving the game by backward induction requires me to start with the councilor's strategies. However, for exposure convenience, assume that I solved the game for the councilors and found the equilibrium cost of governing $C_G(N)$ when there are $N$ city councilors. In the next section, I define $C_G(N)$ for each mechanism using backward induction, and the definitions are well-behaved. In this section, I use this assumption to simplify the exposition.

I now derive the optimal rents ($r$) and public goods provision ($g$) proposed by the mayor. The mayor benefits from public goods provision, as it increases the chances of her reelection. However, she prefers to contribute as minimum as necessary for the public goods and extract the remaining resources as political rents. The mayor's objective is to maximize her expected utility, subject to the municipal budget constraint.

@@

In equilibrium, the budget binds, and the provision of the public goods maximizes:

@@

The first-order condition for an optimal public goods provision makes the marginal benefits to increase rents, in this case, marginal costs of providing the public goods, equals to the marginal benefits of reelection:

@@

\begin{proposition}
The provision of public goods increases with legislature size when the costs of governing decrease with legislature size.
\end{proposition}

\begin{proof}
To prove this result, I need to show that the mayor's expected utility satisfies the increasing differences in $g$ and $N$. This would mean that when increasing the size of the council, the optimal solution $g^*$ would also increase. However, as $N$ increases discretely, I cannot take the cross-partial derivative on $N$ or even use the implicit function theorem. Instead, I use monotone comparative statics to derive these results \citep{milgrom1994monotone}.

The mayor's expected utility satisfy the increasing differences in $g$ and $N$ when, for $g > g'$ and $N+1 > N$,\footnote{Moreover, the requirement that the decision and type space is a complete lattice, is satisfied for $\mathbb{R} \times \mathbb{N}$.} I have:

@@

After substituting, I have:

@@

Multiplying both sides by $\dfrac{1}{g - g'}$ and taking the limit when $g \rightarrow g'$, I have:

@@

It is now clear to see that the function satisfies the increasing differences in $g$ and $N$ when the governing costs of a $N+1$-sized city council are lower than the governing costs of a $N$-sized city council:

@@

This provides the first empirical hypothesis to test: \textit{if the governing costs decrease when the legislature increases, then the public goods provision increase when the legislature size increase}.

At this stage, I need to solve for the existence of these governing costs. Below I show that this proposition is valid for the *partisan* and *hybrid* mechanisms, but it is not suitable for the *non-partisan* mechanism.

@@

In the \textit{non-partisan reversal} mechanism, I assume that when the council rejects the mayor's office, a reversal stage starts with the selection of one councilor randomly. To find the stationary sub-game perfect Nash equilibrium, suppose that there were $k-1$ rejections, and the game is at the $k$-th stage. A councilor accepts the proposer's offer if, and only if, to get the offer at $k$ is better than to wait until the next stage $k+1$. If the offer is $x_i$, then:

@@

On the left-hand side, I place the offer. On the right-hand side, there are two components. The first is the amount that the councilor $i$ gets when she is the proposer. It is equal to the chance that she is recognized as the proposer times the budget in the next round minus the offers she has to make to convince half of the councilors. The second part is the gains if the councilor rejects the current received offer but still gets an offer in the following round. It is equal to the chance that she is not recognized as the proposer times the chance that she receives a transfer times the transfer amount. Note that the councilor's offer as the proposer is the same as the offer she wants to receive. This is because the city councilors are exchangeable. The solution is symmetric for all councilors receiving an offer (this means that, without loss of generality, I could have dropped the $i$ in the solution). After some algebra, the offer $x_i$ has to be greater than or equal to:

@@

Proceeding backwards, at the mayor's proposal stage $k=0$ and the mayor is going to offer $\underline{x}(0, N) = \dfrac{2 \delta R}{2N + 1}$ to half of the councilors. In this context, the governing costs for the mayor are:

@@

\begin{proposition}
In the \textit{non-partisan reversal} mechanism, the governing costs always increase when the size of the legislature increases.
\end{proposition}

@@

The difference in costs when there are $N+1$ city councilors versus when there are $N$ city councilors is equal to:

@@

Therefore, the costs always increase when the size of the legislature increases.

@@

\begin{corollary}
In the \textit{non-partisan reversal} mechanism, public goods provision decreases as the legislature size increases.
\end{corollary}

This result is crucial to understand the role of politics in my model. If the governing costs increase with legislature size, then any increase in public goods provision associated with legislature size has come from a different mechanism. In my case, I show that there is a \textit{political} reason for increasing the public goods provision.

@@

I now look into the other extreme, which is a fully politicized legislature. In this case, when the mayor implements the policy, it generates a value of $p > 0$. A legislator aligned with the mayor receives a benefit $p$ while an opposition legislator receives a cost of $-p$. Moreover, as in the non-partisan reversal, the mayor can choose to distribute $x_i$ private goods for a legislator $i$. In this case, when the policy is implemented, the aligned legislator's utility equals $x_i + p$. When the legislator is not aligned with the mayor, her utility equals $x_i - p$. If the policy is not implemented, then I assume a reversal policy of zero in all choice vectors $g = x_i = p = 0$.

A legislator aligned with the mayor always supports the mayor's policy, regardless of receiving private goods. This is because as $p > 0$, the councilor always receives positive benefits from the policy.\footnote{The mayor could even impose a cost to the city councilors aligned with the government. Although this seems unrealistic, it is mostly the case when government-aligned legislators are recognized to draft legislation for the mayor's office or even take the fallout from a failed policy.} However, politicians not aligned with the mayor require compensation when their votes are needed. In this case, if the mayor's proposal requires the support of an opposition politician, then the mayor compensates the councilor for the policy costs. In such case:

@@

Optimality dictates that the mayor offers $x_i = p$ to the opposition legislator. Let $\gamma$ be the ex-ante probability of electing a politician aligned with the mayor. Then, the costs of governing when the legislature size is $N$ is equal to the expected number of politicians in the opposition that the mayor has to compensate minus the amount needed to implement the policy, in this case, half of them:

@@

\begin{proposition}
In the \textit{partisan reversal} mechanism, if the chance of electing a mayor-aligned politician is greater than one-half, then the costs of govern decrease as the legislature size increases.
\end{proposition}

@@

The differences in costs when there are $N+1$ versus when there are $N$ legislators are equal to:

@@

Therefore, when $\gamma < \dfrac{1}{2}$, the costs of govern increase. Otherwise, the costs decrease.

@@

\begin{corollary}
In the \textit{partisan reversal} mechanism, if the chance of electing a mayor-aligned politician is greater than $1/2$, then public goods provision increase as the legislature size increases.
\end{corollary}

This result shows that when the chances of electing an aligned politician are sufficiently high, the government costs decrease. As a response, the public goods provision (and also rents) increase. The rate distribution between these two vectors is determined by the marginal change in the utility from rents and the utility from reelection. In any case, the amount allocated for both increases.

@@

The \textit{hybrid reversal} mechanism combines both partisan and non-partisan motivations. In the hybrid mechanism, a city councilor aligned with the government favors the mayor's offer if:

@@

An opposition politician, on the other hand, favors the mayor's offer if, and only if:

@@

I can decompose these costs into two components—first, the costs in terms of rents. Second, the costs (or benefits) from political alignment. In this context, the governing costs depend on the mayor's offers for councilors in the mayor's coalition versus the mayor's proposals for councilors in the opposition. The ex-ante chance of an opposition member be elected equal to $1-\gamma$, and taking the weighted averages, these costs become:

@@

\begin{proposition}
In the \textit{hybrid reversal} mechanism, if $\gamma \leq \dfrac{1}{p}\left[ \dfrac{1}{(2N+1)(2N+3)} \right]$, then the costs of govern increase as the legislature size increases. Otherwise, the costs decrease when the legislature size increases.
\end{proposition}

@@

The difference in costs when there are $N+1$ city councilors versus the costs when there are $N$ councilors is equal to:

@@

The conditions for the differences in governing costs decrease when the legislature size increases are:

@@

\begin{corollary}
In the \textit{hybrid reversal} mechanism, when the chances of electing a city councilor aligned with the mayor are sufficiently high, then increasing council size lowers the bargaining costs for mayors.
\end{corollary}

The threshold $\underline{\gamma}$ for the probability of electing a politician aligned with the government is decreasing in the value of the mayor's policy $p$ and the legislature size $N$. This means that higher political values and larger legislatures make it easier to satisfy the electoral threshold.

@@

Our model provides two empirically testable hypotheses. Consider a city council comprised of $N$ legislators that increase its council size to $N+1$. Then:

\begin{enumerate}
  \item[H1. ] The governing costs decrease with legislature size when the chance of electing a government-aligned legislator is sufficiently high.
  \item[H2. ] The provision of public goods increase when the governing costs decrease with legislature size.
\end{enumerate}

In the paper, I show that the provision of education and healthcare improves in Brazil. This results from the fact that the chances of electing a city councilor aligned with the mayor are 91%.

@@

I use three information sources, either from Brazilian governmental agencies or available online. Table \ref{varsources} displays the primary sources and their respective URLs.

\begin{sidewaystable}[!htbp] \centering 
\caption{Data Sources} 
\label{varsources}
\scalebox{0.8}{
  \begin{tabular}{@{\extracolsep{5pt}}llll} 
  \\[-1.8ex]\hline\hline \\[-1.8ex] 
  Source Code & \multicolumn{1}{c}{Source Name} & \multicolumn{1}{c}{Description} & \multicolumn{1}{c}{URL} \\ 
  \hline \\[-1.8ex] 
  & & & \\
  DataSUS & Brazilian Health Ministry Data Service & Collects data on Health Care & \url{http://www.datasus.gov.br} \\
  & & & \\
  IBGE & Brazilian Institute of Geography and Statistics & Collects data Geography, Economics and Demography & \url{http://www.ibge.gov.br} \\ 
  & & & \\
  INEP & Data Service of Ministry of Education & Collects data on performance of Education & \url{http://www.inep.gov.br} \\ 
  & & & \\
  InterLegis & Senate Legislative Data Service & Collect Data on Legislative in Brazil & \url{http://www.interlegis.leg.br} \\
  & & & \\
  IPEA & Brazilian Institute of Applied Economics & Collects data on the Economy & \url{http://www.ipeadata.gov.br} \\ 
  & & & \\
  MDS & Social Security Ministry & Collect data on Social Coverage and Effectiveness of Social Programs & \url{http://www.mds.gov.br} \\ 
  & & & \\
  TSE & Superior Electoral Tribunal & Collects data on elections results & \url{http://www.tse.jus.br} \\
  & & & \\
  \hline\hline \\[-1.8ex] 
  \end{tabular} 
}
\end{sidewaystable}

@@

I coded the legislation approved by the councilors in 63 municipalities within ten thousand inhabitants away from the council size thresholds. I classified the legislation into five categories:

\begin{itemize}
\item \textbf{Public Goods}: Legislation that provides public goods or services.
\item \textbf{Oversight}: Legislation that requests information to the mayor's office or the bureaucracy about the provision of services.
\item \textbf{Others}: Legislation that is not classified as Local Public Goods, Public Goods, or Oversight. Usually honors or procedures.
\item \textbf{Education and Health Care}: Legislation about Education or Health Care provision.
\end{itemize}

@@

The validity of the causal claim relies on three assumptions:

>>>>

[^danilo]: Senior Lecturer, School of Social and Political Sciences, University of Lincoln, <dfreire@lincoln.ac.uk>, <https://danilofreire.github.io>.

@@

\appendix
\addtocounter{section}{1}

@@

### Model Primitives

Consider an strategic interaction between a mayor $M$ and $N \geq 3$ city councilors indexed as $i \in \{1, 2, \cdots, N \}$. The mayor is the head of the local executive branch and submits policy proposals to the city councilors. Policy proposals comprise funds allocated to public goods provision ($g$), rents for the mayor ($r$), and transfers to the city councilors (\textbf{x}).

The city council votes the mayor's proposal, which is implemented if they accept it. Otherwise, a reversal policy takes place, indicating that the mayor failed to secure the majority of the councilors' votes. In this case, the policy proposal either stops or is transferred to the council. The different types of reversal policies are crucial for our argument, as they change the bargaining position of the councilors relative to the mayor. We investigate three types of reversal policies: 1) a \textit{non-partisan} reversal mechanism, in which city councilors have no partisan affiliation; 2) a \textit{partisan} reversal mechanism, where the council is divided into supporters and non-supporters of the mayor; 3) and finally, a \textit{hybrid} reversal mechanism, which combines partisan and non-partisan features. The main paper focuses on the hybrid model, as it is the more realistic scenario. However, comparing the partisan and non-partisan reversal mechanisms provides further evidence that the partisan mechanism drives our results.

The mayor needs to secure the support of at least half of the city councilors to implement his or her policy. Similar to \citet{buenodemesquita2005logic}, we assume that the selectorate is composed by the $N$ city councilors. That is, these are the individuals who have a say in the mayor's policy proposals. The winning coalition, which is the minimum number of members required to implement the policy, equals half of the selectorate ($N/2$). City councilors are motivated by policy concerns $p$ and transfers received either from the mayor or from a councilor \textbf{x} who proposes the policy. We leave the content of these transfers open, as this allows us to analyze how mayors combine transfers with other incentives to pressure city councilors. For instance, a mayor may grant councilors electoral favors, spend in areas where they have political or personal interests, offer portfolio or public sector jobs to associates, or even provide campaign contributions and bribes. These transfers affect the councilors' decisions and generate governing costs $C_G$. Governing costs vary according to the type of reversal mechanism and the size of the legislature.

Finally, after paying the aforementioned governing costs, the mayor proposes some level of public goods provision $g$ and rents $r$. Public goods provision helps the mayor to get reelected. Rents are for the mayor's direct benefit and do not contribute to his or her electoral success. This makes the mayor's expected utility a sum of the gains from rents plus the benefits from reelection. Both the utilities from rents $u(.)$ and the probability of reelection $\pi(.)$ are concave functions, meaning that more rents or more public goods increase his or her utility at a decreasing rate.\footnote{In technical terms, $u'>0$ and $\pi'>0$; and $u''<0$ and $\pi''<0$.} The probability of reelection is multiplied by the benefit from holding office $B_M > 0$. This benefit captures the tangible and intangible gains the mayor receives from holding the public office. We assume these benefits are high enough to rule out an equilibrium with only rents and transfers, which is equivalent to saying that mayors strongly prefer to be reelected. If the mayor's policy proposal is approved, then the expected utility for the mayor is:

@@

The mayor's policy choices are subject to municipal budget constraints. The municipality has $R > 0$ resources and cannot run deficits, that is, it must always balance its budget. We also assume that there are enough resources for the mayor to govern.

Let $C_G$ be the mayor's governing costs. The budget balance constraints implies that mayors will only offer public services and rents that satisfy following inequality:

@@

The expected utility for the city councilors depends on which reversal policy takes place. As mentioned above, we explore the implications of three different mechanisms. First, in the \textit{non-partisan} reversal mechanism, when the council rejects the mayor's proposal, a reversal stage starts with resources diminishing by a factor $\delta \in (0,1)$ and with the random selection of one councilor. The randomly selected councilor becomes the new proposer. If her proposal is accepted, it is implemented. If rejected, the budget shrinks again by a factor $\delta$, and another city councilor is chosen to propose the policy. The process continues until one proposal is finally accepted.

In the second reversal mode, the \textit{partisan} reversal mechanism, we assume that each councilor has a party affiliation. Party affiliations are mutually exclusive, that is, one councilor cannot belong to two parties at once. If a city councilor is affiliated to a party aligned with the mayor, we say that she belongs to the government coalition $G \subset \{1, 2, \cdots, N \}$. Otherwise, the city councilor belongs to the opposition $O = \{1, 2, \cdots, N \} \setminus G$. The mayor-sponsored policy generates a political value of $p > 0$ for the councilors, and whether this adds or reduces the councilor's utility depends on his or her party affiliation. If the mayor's proposal is rejected, a city councilor aligned with the mayor receives zero instead of $p$. In contrast, a city councilor from the opposition receives zero if the proposal is rejected and $-p$ if is is enacted. Although these assumptions are rather strong, they capture the underlying political dynamics of political partisanship, in which councilors and mayors in the same coalition share similar policy preferences. Voting against the mayor's proposals defies the partisan logic.

Lastly, in the \textit{hybrid} reversal mechanism, we combine partisanship with legislature bargaining costs. Here councilors receive a partisan utility from the mayor's policy, but if they reject the policy, a reversal policy similar to the non-partisan mechanism takes place. In the mayor's proposal stage, city councilors in the mayor's coalition receive a benefit of $p > 0$ when the mayor's policy is implemented. Opposition members receive $-p$ if the mayor's proposal is implemented. In case the mayor's proposal is rejected, as the city council does not control the executive when the two branches are separated, there is no provision of public goods ($p=0$). Moreover, in equilibrium, we assume that city councilors prefer to receive transfers instead of providing public goods. This is in line with the empirical evidence from most developing economies [e.g., @FerrazFinan2011; @Nichter2010].

@@

\begin{enumerate}
  \item The mayor learns how many government $|G|$ and opposition $|O|$ legislators have been elected.
  \item The mayor proposes a policy vector $(r, g, x)_M$.
  \item The city council votes the proposal.
  \begin{itemize}
    \item If the council accepts the proposal, it is implemented and the game ends.
    \item Otherwise, a reversal policy is implemented.
  \end{itemize}
  \item (Reversal Policy) We propose three reversal policies, namely \textit{partisan}, \textit{non-partisan}, and \textit{hybrid}. We have introduced them above and will discuss them in further detail in the following sections.
\end{enumerate}

The solution concept we use in this model is the subgame perfect Nash equilibrium. This game requires that the strategies follow a Nash equilibrium in each subgame. We find the solution by backward induction. In the model with infinitely repeated proposals, we extend the equilibrium concept and require that the equilibrium is stationary. This implies that if the politician accepts an offer at time $k+1$, they should take the same offer at time $k$. This assumption holds at each point of the game. Stationarity provides a method to find a proposal that the politician accepts at any stage in the game. In our case, the optimal offer for the mayor is at $k=0$, which indicates an immediate policy implementation.

\begin{definition}
The game equilibrium consists of a sequence of strategic policy vectors $(g^*, r^*, x^*)_k$, indexed in $k \in \{0, 1, \cdots, \infty\}$, such that the mayor and the councilors have no profitable deviations from the equilibrium at any stage of the game.
\end{definition}

Moreover, we assume that players have no dominated strategies, even if the policy is not implemented due to insufficient support.

@@

Solving the game by backward induction requires us to know which strategies councilors may adopt. However, assume for now that we have solved the game for the councilors and found the equilibrium cost of governing $C_G(N)$ when there are $N$ city councilors. In the next section, we define $C_G(N)$ for each mechanism using backward induction and show that the definitions are well-behaved. Here we start from this assumption just to simplify our exposition.

We derive the optimal rents ($r$) and public goods provision ($g$) a mayor should propose. Mayors benefits from public goods provision, as it increases their chances of reelection. However, they would rather provide the lowest necessary amount of public goods and extract the remaining resources as political rents. As we noted previously, their objective is to maximize their expected utility, subject to the municipal budget constraint.

@@

In equilibrium, the budget binds, and the provision of the public goods maximizes to:

@@

The first-order condition for an optimal public goods provision makes the marginal benefits of increasing rents---in this case, the marginal costs of providing the public goods---equal to the marginal benefits of reelection:

@@

\begin{proposition}
Public goods provision increases with legislature size if governing costs decrease with legislature size.
\end{proposition}

\begin{proof}
To prove this result, we need to show that the mayor's expected utility satisfies the increasing differences in $g$ and $N$. This means that, when increasing the size of the council, the optimal solution $g^*$ should also increase. However, as $N$ increases discretely, we cannot take the cross-partial derivative on $N$ or use the implicit function theorem. Instead, we use monotone comparative statics to derive these results \citep{milgrom1994monotone}.

The mayor's expected utility satifies the increasing differences in $g$ and $N$ when, for $g > g'$ and $N+1 > N$,\footnote{The requirement that the decision and type space are a complete lattice is satisfied for $\mathbb{R} \times \mathbb{N}$.} we have:

@@

After some substitutions, we have:

@@

By multiplying both sides by $\dfrac{1}{g - g'}$ and taking the limit when $g \rightarrow g'$, we have:

@@

It is now clear that the function satisfies the increasing differences in $g$ and $N$ when the governing costs of a $N+1$-sized city council are lower than the governing costs of a $N$-sized city council:

@@

This provides the first empirical hypothesis we test in our article: \textit{public goods provision increases along with legislature size, but only if governing costs decrease when legislature size increases}.

At this stage, we need to solve for the existence of these governing costs. In the following sections we show that while this proposition is valid for the *partisan* and *hybrid* mechanisms, it is not applicable to the *non-partisan* mechanism.

@@

In the \textit{non-partisan reversal} mechanism, we assume that if the council rejects the mayor's proposal, a reversal policy starts with the random selection of one councilor. To find the stationary subgame perfect Nash equilibrium, suppose that there were $k-1$ rejections and the game is at the $k$-th stage. A councilor accepts the proposer's offer if, and only if, getting the offer at $k$ is better than waiting until the next stage $k+1$. If the offer is $x_i$, then:

@@

We place the offer on the left-hand side. There are two components on the right-hand side. The first is the amount that the councilor $i$ receives when he/she is the proposer. The amount is equal to the chance that they are recognized as the proposer, times the budget in the next round, minus the offers they have to make to convince half of the councilors. The second part denotes the gains obtained by the councilor if he/she rejects the current offer but still receives an offer in the following round. This is equal to the chance that they are not recognized as the proposer, times the chance that they receive a transfer, times the transfer amount. Note that the councilor's offer as the proposer is the same as the offer they want to receive, as city councilors are exchangeable. The solution is symmetric for all councilors receiving an offer (this means that, without loss of generality, we could have dropped the $i$ in the solution). After some calculations, the offer $x_i$ has to be greater than or equal to:

@@

Proceeding backwards, at the mayor's proposal stage $k=0$, the mayor offers $\underline{x}(0, N) = \dfrac{2 \delta R}{2N + 1}$ to half of the councilors. In this context, the governing costs for the mayor are:

@@

\begin{proposition}
In the \textit{non-partisan reversal} mechanism, governing costs always increase when legislature size increases.
\end{proposition}

@@

Comparing city councils of size $N$ with those of size $N+1$, we find that the difference is costs is:

@@

Therefore, costs always increase when legislature size increases.

@@

\begin{corollary}
In the \textit{non-partisan reversal} mechanism, public goods provision decreases as legislature size increases.
\end{corollary}

This result is crucial for us to explain the role of politics in our model. If governing costs increase with legislature size, then any increase in public goods provision associated with legislature size comes from a different mechanism. In this case, we show that there is a \textit{political} reason to increase the provision of public goods.

@@

We now focus on the other extreme, which is a fully politicized legislature. In this case, when the mayor implements the policy, it generates a value of $p > 0$. A legislator aligned with the mayor receives a benefit $p$ while an opposition legislator receives a cost of $-p$. Moreover, as in the non-partisan reversal, the mayor may choose to distribute $x_i$ private goods for a legislator $i$. In this case, when the policy is implemented, the aligned legislator's utility equals $x_i + p$. When the legislator is not aligned with the mayor, their utility equals $x_i - p$. If the policy is not implemented, then we assume a reversal policy of zero in all choice vectors $g = x_i = p = 0$.

A legislator aligned with the mayor always supports the mayor's policy, regardless of receiving private goods. This is because $p > 0$, that is, the councilor always receives positive benefits from the policy.\footnote{The mayor could even impose costs to aligned city councilors. Although this seems unrealistic, it is mostly the case when government-aligned legislators are called to draft legislation for the mayor's office or to take the fallout from a failed policy.} However, politicians who not aligned with the mayor require compensation when their votes are needed. In this case, if the mayor's proposal requires the support of an opposition politician, then the mayor compensates the councilor for the policy costs. In such case:

@@

Optimality dictates that the mayor offers $x_i = p$ to the opposition legislator. Let $\gamma$ be the ex-ante probability of electing a politician aligned with the mayor. Then, the governing costs when the legislature size is $N$ equals the expected number of opposition members the mayor has to compensate, minus the amount needed to implement the policy. In this case, it is half of the city council:

@@

\begin{proposition}
In the \textit{partisan reversal} mechanism, if the chances of electing a mayor-aligned politician are greater than $1/2$, then governing costs decrease as legislature size increases.
\end{proposition}

@@

When we compare the difference in costs between $N$ and $N+1$ legislators, we find that:

@@

Therefore, when $\gamma < \dfrac{1}{2}$, governing costs increase. Otherwise, such costs decrease.

@@

\begin{corollary}
In the \textit{partisan reversal} mechanism, if the chances of electing a mayor-aligned politician is greater than $1/2$, then public goods provision increase as legislature size increases.
\end{corollary}

This result shows that, when the likelihood of electing an aligned politician is sufficiently high, governing costs tend to decrease. Consequently, the provision of public goods tends increases, as well as the rents. The marginal change in the utility from rents and the utility from reelection determines the allocation between these two vectors. In any case, the amount distributed to both increases.

@@

The \textit{hybrid reversal} mechanism combines both partisan and non-partisan motivations. Here, a city councilor aligned with the government favors the mayor's offer if:

@@

An opposition politician, in contrast, favors the mayor's offer if, and only if:

@@

We divide these costs into two components. First, the costs in terms of rents. Second, the costs (or benefits) from political alignment. In this context, governing costs depend on what the mayor offers to their supporters in the council versus what they provide to members of the opposition. The ex-ante chance of an opposition member being elected is equal to $1-\gamma$, therefore, after taking the weighted averages, these costs become:

@@

\begin{proposition}
In the \textit{hybrid reversal} mechanism, if $\gamma \leq \dfrac{1}{p}\left[ \dfrac{1}{(2N+1)(2N+3)} \right]$, then governing costs increase as legislature size increases. Otherwise, such costs decrease when legislature size increases.
\end{proposition}

@@

When we compare the costs of $N$ versus $N+1$ legislators, we find that:

@@

The conditions for the difference in governing costs to decrease when legislature size increases are as follows:

@@

\begin{corollary}
In the \textit{hybrid reversal} mechanism, when the chances of electing a mayor-aligned city councilor are sufficiently high, then increasing council size lowers the bargaining costs for mayors.
\end{corollary}

The threshold $\underline{\gamma}$ for the probability of electing an aligned politician decreases according to the value of the mayor's policy $p$ and the legislature size $N$. Higher political values and larger legislatures make it easier to satisfy the electoral threshold.

@@

Our model provides two empirically testable hypotheses. Consider a city council comprised of $N$ legislators that increase its size to $N+1$. Then:

\begin{enumerate}
  \item[H1. ] Governing costs decrease with legislature size when the chances of electing a government-aligned legislator are sufficiently high.
  \item[H2. ] Public goods provision increases if governing costs decrease with legislature size.
\end{enumerate}

In the main paper, we show that larger legislatures improve the provision of public education and health care in Brazil. What drives this result is the 91% chance of electing a mayor-aligned city councilor to the local chamber.

@@

We use three data sources in our work, mainly from Brazilian governmental agencies. Table \ref{tab:varsources} displays the primary sources we use here and their respective URLs.

\begin{table}[!htbp] \centering
\small
\caption{Data Sources}
\label{tab:varsources}
\scalebox{0.8}{
  \begin{tabular}{@{\extracolsep{5pt}}llll}
  \\[-1.8ex]\hline\hline \\[-1.8ex]
  Source Code & \multicolumn{1}{c}{Source Name} & \multicolumn{1}{c}{Data Provided} & \multicolumn{1}{c}{URL} \\
  \hline \\[-1.8ex]
  & & & \\
  DataSUS & Brazilian Health Ministry Data Service & Health care data & \url{http://www.datasus.gov.br} \\
  & & & \\
  IBGE & Brazilian Institute of Geography and Statistics & Geographic, economic, and demographic data & \url{http://www.ibge.gov.br} \\
  & & & \\
  INEP & Ministry of Education Data Service & Education performance & \url{http://www.inep.gov.br} \\
  & & & \\
  InterLegis & Senate Legislative Data Service & Legislative data & \url{http://www.interlegis.leg.br} \\
  & & & \\
  IPEA & Brazilian Institute of Applied Economics & Economic data & \url{http://www.ipeadata.gov.br} \\
  & & & \\
  MDS & Social Security Ministry & Coverage of social programs & \url{http://www.mds.gov.br} \\
  & & & \\
  TSE & Supreme Electoral Court & Electoral data & \url{http://www.tse.jus.br} \\
  & & & \\
  \hline\hline \\[-1.8ex]
  \end{tabular}
}
\end{table}
\normalsize

@@

We code the legislation approved by the councilors of 63 municipalities around a ten thousand inhabitants buffer from the council size thresholds. We classify the legislation into four categories:

\begin{itemize}
\item \textbf{Public Goods}: Legislation that provides public goods or services.
\item \textbf{Oversight}: Legislation that requests information to the mayor's office or the bureaucracy about the provision of services.
\item \textbf{Education and Health Care}: Legislation about education or health care provision.
\item \textbf{Others}: Legislation that is not classified as local public goods, public goods, or oversight. Usually honors or procedures.
\end{itemize}

@@

The validity of our causal claim relies on three assumptions:
